Supplements for ``Tensor denoising and completion based on ordinal observations''

### 2020-01-21 ICLM/miaoyan/ordinalT_supp.tex
update draft

\begin{lem}~\label{lem:KL}
Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$. Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$. Define
\begin{equation}\label{eq:ass}
A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(\theta+b_\ell)-f(\theta+b_{\ell-1})\right].
\end{equation}
Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
\[
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{lem}
\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
\[
\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(\theta_\omega+b_{\ell})-f(\theta_\omega+b_{\ell-1}),
\]
for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$. For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$). Based on Lemma~\ref{lem:KLentry} and Taylor expansion, 
\begin{align}
\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(\theta+b_{\ell})-f(\theta+b_{\ell-1})-f(\theta^*+b_{\ell})+f(\theta^*+b_{\ell-1})\right]^2\over f(\theta^*+b_{\ell})-f(\theta^*+b_{\ell-1})}\\
 &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(\eta_\ell+b_\ell)-\dot{f}(\eta_{\ell-1}+b_{\ell-1})\right]^2 \over f(\theta^*+b_\ell)-f(\theta^*+b_{\ell-1})} \left(\theta-\theta^*\right)^2+{\dot{f}^2(\eta_1+b_1) \over f(\theta^*+b_1)} (\theta-\theta^*)^2\\
 &\quad \quad\quad \quad  +{\dot{f}^2(\eta_{L-1}+b_{L-1})\over 1-f(\theta^*+b_{L-1})}(\theta-\theta^*)^2,
\end{align}
where $\eta_\ell, \eta_{\ell-1}$ fall between $\theta$ and $\theta^*$, and $\check b_{\ell}$ falls between $b_{\ell-1}$ and $b_\ell$. Therefore,
\begin{equation}\label{eq:entrywise}
\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
\end{equation}
where we have used the bound~\eqref{eq:ass} and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function. Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
\[
\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{proof}

>>>>

\begin{lem}~\label{lem:KL}
Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$. Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$. Define
\begin{equation}\label{eq:ass}
A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(\theta+b_\ell)-f(\theta+b_{\ell-1})\right].
\end{equation}
Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
\[
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{lem}
\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
\[
\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(\theta_\omega+b_{\ell})-f(\theta_\omega+b_{\ell-1}),
\]
for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$. For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$). Based on Lemma~\ref{lem:KLentry} and Taylor expansion, 
\begin{align}
\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(\theta+b_{\ell})-f(\theta+b_{\ell-1})-f(\theta^*+b_{\ell})+f(\theta^*+b_{\ell-1})\right]^2\over f(\theta^*+b_{\ell})-f(\theta^*+b_{\ell-1})}\\
 &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(\eta_\ell+b_\ell)-\dot{f}(\eta_{\ell-1}+b_{\ell-1})\right]^2 \over f(\theta^*+b_\ell)-f(\theta^*+b_{\ell-1})} \left(\theta-\theta^*\right)^2+{\dot{f}^2(\eta_1+b_1) \over f(\theta^*+b_1)} (\theta-\theta^*)^2\\
 &\quad \quad\quad \quad  +{\dot{f}^2(\eta_{L-1}+b_{L-1})\over 1-f(\theta^*+b_{L-1})}(\theta-\theta^*)^2,
\end{align}
where $\eta_\ell, \eta_{\ell-1}$ fall between $\theta$ and $\theta^*$, and $\check b_{\ell}$ falls between $b_{\ell-1}$ and $b_\ell$. Therefore,
\begin{equation}\label{eq:entrywise}
\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
\end{equation}
where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function. Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
\[
\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{proof}
